Optimal (Multiway) Spatial Joins

### 2024-05-24 msj.tex
yufei

\begin{lemma} \label{lmm:hv:type1:proj}
    $|\J(R_1', R'_2, ..., R'_{k-2})| \le \out$.
\end{lemma}

We remind the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.

@@

\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r' \in R_i'$, define
\myeqn{
    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): r' = \gbot(\bm{t})} \ytop(B_\bm{t}). 
}
The value $\maxtop(r')$ is set to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$. When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
\myeqn{
    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
}
and call it the {\em top-sliced rectangle} of $r'$.

@@

We pay particular attention to the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$, defined in \eqref{eqn:cross}, is the set of segments in $H$ crossing $r^*$. The following lemma, proven in Appendix~\ref{app:hv}, presents some crucial properties for these sets.

@@

        \item Consider any $k$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$. Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$. Then, it must hold that $\dcross_H(\bm{t}) \subseteq \cross_H(r^*)$. Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ have been sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
    }
\end{lemma}

@@

Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$. Statement (1) of the above lemma assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$. Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.

\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}. First, identify the rectangle $r' = \gbot(\bm{t})$ and fetch its (already computed) top-sliced rectangle $r^*$. Second, obtain $B_\bm{t}$ and $\gleft(\bm{t})$. These two steps require $O(k)$ time. Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates. For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely,  whether $h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$; this can be done in constant time. Abort the scan as soon as $h \notin \dcross_H(\bm{t})$. This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|)$ time. Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + |\J^*|) = O(k \cdot \out)$ time.

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:proj}.}

@@

>>>>

\begin{lemma} \label{lmm:hv:type1:recur-output}
    $|\J(R_1', R'_2, ..., R'_{k-2})| \le \out$.
\end{lemma}

We remind the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:recur-output}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.

@@

\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r' \in R_i'$, define
\myeqn{
    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): \gbot(\bm{t}) = r'} \ytop(B_\bm{t}). 
}
We set $\maxtop(r')$ to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$. When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
\myeqn{
    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
}
and call it the {\em top-sliced rectangle} of $r'$.

@@

We are particularly concerned with the sets $\cross_H(r^*)$ of the rectangles $r^*$ in $R^*_i$, where $\cross_H(r^*)$ --- defined in \eqref{eqn:cross} --- is the set of segments in $H$ crossing $r^*$. The following lemma, proven in Appendix~\ref{app:hv}, presents some useful properties for these sets.

@@

        \item Consider any $k$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$. Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$. Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$. Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
    }
\end{lemma}

@@

Finding the $\cross_H(r^*)$ sets of all $r^* \in R^*_i$ is an instance of the find-all-sorted version of Problem $\mathscr{D}$. Statement (1) of Lemma~\ref{lmm:hv:type1:cross-r*} assures us that the total time to do so for all $R^*_1, ..., R^*_{k-2}$ is bounded by $O(n \log n + \out)$. Note that, for each $\cross_H(r^*)$ computed, the (horizontal) segments therein has been sorted in ascending order of y-coordinate.

\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}. First, identify the rectangle $r' = \gbot(\bm{t})$ and fetch its (already computed) top-sliced rectangle $r^*$. Second, obtain $B_\bm{t}$ and $\gleft(\bm{t})$. These two steps require $O(k)$ time. Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates. For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely,  whether $h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$; this can be done in constant time. Abort the scan as soon as $h \notin \dcross_H(\bm{t})$. This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|)$ time. Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)| = O(k \cdot \out)$ time, where the last step used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:recur-output}.}

@@

\extraspacing {\bf Proof of Lemma~\ref{lmm:hv:type1:cross-r*}.}
